SSA based global code motion of identical computations

### 2017-04-17 paper.tex
fixes to section 1

This paper describes code-motion of congruent instructions \cite{briggs1997}, a
technique that uses the information computed for \PRE{} to detect identical
computations but has a transformation phase whose goal differs from \PRE{}: it
removes identical computations from different branches of execution.  These
identical computations in different branches of execution are not redundant
computations at run-time and the number of run-time computations is not
reduced. It is not a redundancy elimination pass, and thus it has different cost
function and heuristics than \PRE{} or CSE. It is more similar to global code
scheduling \cite{dragonbook,click1995global} in the sense that it will only move
computations. Code hoisting can reduce the critical path length of execution in
out of order machines. As more instructios are available at the hoisting point,
the hardware has more instructions to reorder. Following example illustrates how
hoisting can improve performance by exposing more ILP.

@@

In this program the computations of tmax and tmin are identical to the
computations of tmin and tmax of sibling branch respectively. Both tmax and tmin
depends on inv which depends on a division operation which is generally more
expensive than the addition, subtraction and multiplication operations. The
total latency of computation across each branch is:
$C_{div} + 2(C_{sub} + C_{mul})$
Or, for out of order processors with two add units and two multiply units:
$C_{div} + C_{sub} + C_{mul}$

Now if the computation of tmax and tmin are hoisted outside the
conditionals, the C code version would look like this:
\begin{verbatim}
float fun(float d, float min, float max, float a) {
  float tmin, tmax, tmin1, tmax1, inv;

  tmin1 = (min - a);
  tmax1 = (max - a);

  inv = 1.0f / d;
  tmin1 = tmin1 * inv;
  tmax1 = tmax1 * inv;

  if (inv >= 0) {
    tmin = tmin1;
    tmax = tmax1;
  } else {
    tmin = tmax1;
    tmax = tmin1;
  }

  return tmax + tmin;
}

>>>>

This paper describes code-motion of congruent instructions \cite{briggs1997}, a
technique that uses the information computed for \PRE{} to detect identical
computations but has a transformation phase whose goal differs from \PRE{}: it
removes identical computations from different branches of execution.  These
identical computations in different branches of execution are not redundant
computations at run-time and the number of run-time computations is not
reduced. It is not a redundancy elimination pass, and thus it has different cost
function and heuristics than \PRE{} or CSE. It is more similar to global code
scheduling \cite{dragonbook,click1995global} in the sense that it will only move
computations. Code hoisting can reduce the critical path length of execution in
out-of-order machines. As more instructions are available at the hoisting point,
the hardware has more independent instructions to reorder. The following example
illustrates how hoisting can improve performance by exposing more ILP.

@@

In this program the computations of tmax and tmin are identical to the
computations of tmin and tmax of sibling branch respectively. Both tmax and tmin
depend on inv which depends on a division operation which is generally more
expensive than the addition, subtraction, and multiplication operations. The
total latency of computation across each branch is: $C_{div} + 2(C_{sub} +
C_{mul})$.  For an out-of-order processor with two add units and two multiply
units, the total latency of computation is $C_{div} + C_{sub} + C_{mul}$.

Now if the computation of tmax and tmin are hoisted outside the
conditionals, the C code version would look like this:
\begin{verbatim}
float fun(float d, float min, float max, float a) {
  float tmin, tmax;
  float inv = 1.0f / d;
  float x = (min - a) * inv;
  float y = (max - a) * inv;
  if (inv >= 0) { tmin = x; tmax = y; }
  else          { tmin = y; tmax = x; }
  return tmax + tmin;
}
